---
title: "Nutrition_data visualization"
author: "Racheal_wanyoike"
date: '2022-06-26'
output:
  html_document:
    theme:
      bootstrap: journal
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Obtained data

```{r,echo=TRUE}
nutrition_data <- read.csv ("https://raw.githubusercontent.com/fivethirtyeight/data/master/nutrition-studies/raw_anonymized_data.csv")

```

Next, I cleaned the data by selecting 12 variables that may have an association to cancer. 

```{r nutrition_data, include=TRUE}
library(dplyr)
cancerlinktofoods <- select(nutrition_data,cancer,BEEFPORKDISHFREQ,DRIEDFRUITFREQ,OTHERFRESHFRUITFREQ,CORNFREQ,SUGARINCOFFEE,SALTFREQ,HOTTEAFREQ,RAWTOMATOESFREQ,MILKFREQ,BUTTERFREQ,OTHERBREADSFREQ, BROCCOLIFREQ)
cancerlinktofoods
```

To understand the numeric variables and  in my cancerlinktofoods dataset, I plotted a summary box plot.

```{r,echo=TRUE}
cancerlinktofoods$cancer = factor(cancerlinktofoods$cancer)
summary(cancerlinktofoods)
boxplot(cancerlinktofoods$BEEFPORKDISHFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$DRIEDFRUITFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$OTHERFRESHFRUITFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$CORNFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$SUGARINCOFFEE~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$HOTTEAFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$RAWTOMATOESFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$BUTTERFREQ~cancerlinktofoods$cancer)
boxplot(cancerlinktofoods$BROCCOLIFREQ~cancerlinktofoods$cancer)

```

The boxplot chart helps visualize the summary statistics grouped by the factor(cancer).We can also see outliers.


The percentage of respondents with cancer and those without:
```{r,echo=TRUE}
cancertb <- table(cancerlinktofoods$cancer)
cancerprop <- prop.table(cancertb)
cancerdf <- as.data.frame(cancerprop)
names(cancerdf) <- c("cancer", "Frequency")

library(ggplot2)
ggplot(data=cancerdf, mapping=aes(x=cancer, y=Frequency)) + 
  geom_col(fill="darkgreen", alpha=0.5) +
  scale_y_continuous(labels = scales::percent)
```

